---
title: "Recalling Words"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---

<script type="text/javascript">
 function showhide(id) {
    var e = document.getElementById(id);
    e.style.display = (e.style.display == 'block') ? 'none' : 'block';
 }
</script>

```{r, include=FALSE}
library(mosaic)
library(car)
library(DT) #You may need to run: install.packages("DT") 
library(pander)
Rtypes <- subset(Friendly, condition != "SFR")
Rtypes$condition <- factor(Rtypes$condition, levels = c("Before", "Meshed"))
```

<br />

#### Background

Many teachers and other educators are interested in understanding how to best deliver new content to students. In general, they have two choices of how to do this.

1. The Meshed Approach
    * Deliver new content while simultaneously reviewing previously understood content.

2. The Before Approach
    * Deliver new content after fully reviewing previously understood content.

A study was performed to determine whether the *Meshed* or *Before* approaches to delivering content had any positive benefits on memory recall.

<div style="padding-left:15px;">

##### <a href="javascript:showhide('uniquename')">The Experiment <span style="font-size:8pt;">(click to view)</span></a>


<div id="uniquename" style="display:none;">

Individuals were seated at a computer and shown a list of words. Words appeared on the screen one at a time, for two seconds each, until all words had been shown (40 total). After all words were shown, they were required to perform a few two-digit mathematical additions (like 15 + 25) for 15 seconds to avoid immediate memory recall of the words. They were then asked to write down as many of the 40 words as they could remember. They were given a maximum of 5.3 minutes to recall words.

The process of showing words and recalling words was repeated four times with the same list of words each time (four chances to get it right). The presentation of the first trial was the same for all treatment conditions. However, trials 2, 3, and 4 were slightly different for each treatment condition.

<div style="padding-left:15px;">

The `SFR` group (the control group) stands for Standard Free Recall. In all four trials the same list of 40 words was presented, in a random order each time.

The `Before` group also used the same 40 words during each trial. However, any words that were correctly recalled in a previous trial were presented first, or *before* the words that were not recalled in the last trial. After all the correct words were presented in random order, the non-recalled words were presented in a random order.

The `Meshed` group also used the same 40 words during each trial. However, words that were correctly recalled in a previous trial were alternated with a missed word during the next presentation order. 

</div>

The data records the number of correctly recalled words (out of the 40 possible) from the fourth trial. Results were obtained for 30 students, 10 in each of the three treatment groups: `SFR`, `Before`, and `Meshed`. 

</div>

##### <a href="javascript:showhide('uniquename2')">The Data <span style="font-size:8pt;">(click to view)</span></a>

<div id="uniquename2" style="display:none;">

The results from the study can be found in the `Friendly` data set in R after loading `library(car)`. Press the code button to view.   


```{r}
datatable(Friendly)
```


</div>
</div>

<br />

------------------

#### Hypotheses 

The null hypothesis is that there is not a significant difference in medians between the two methods, Meshed and Before, and the alternative is that there is a difference between them. Formally written as,


$H_0:$ difference in medians $= 0$ 

$H_a:$ difference in medians $\neq 0$

with $\alpha = 0.05$

-------------------

####Analysis

To check the distribution of our data a box plot was shown below. 

```{r}
boxplot(correct ~ condition, data = Rtypes, main = "Reviewing Methods", ylab = "Number of Correct Words Recalled (max 40)", col = c("firebrick2", "green3")) 
```

We can see that there is a slight difference in distribution and medians between Before and Meshed while SRF, the control, is vastly different. 

The medians are 

|Method|Before|Meshed|
|------|------|------|
|Median|39.0  |36.5  |

After looking closer at the data we also observed that both data sets have a lot of ties. A table below shows the numbers and the number of ties in the data set. (note that the number of ties are only from the Meshed and Before data. SRF was not considered because the null is only comparing the medians of Before to Meshed)

|Observed Value | 24 | 30 | 34 | 36 | 37 | 38 | 39 | 40 |
|---------------|----|----|----|----|----|----|----|----|
|Number of Ties | 1  | 2  | 1  | 3  | 2  | 2  | 4  | 5  |

To compensate for the amount of ties in the data a continuity correction was done and the results are as follows. 

```{r, warning = FALSE}
pander(wilcox.test(correct ~ condition, data=Rtypes, mu=0, alternative="two.sided", conf.level=0.95))
```

The test statistic was $W = 62$ which gave a p-value of 0.378. 
Comparing our $p$-value to $\alpha$ we get:
$$
(p = 0.378 > \alpha)
$$
Therefore, we have insufficient evidence to reject the null hypothesis.  

-------------------

#### Interpretation 

The results of the study shows a lack of evidence to favor anyone method over the other. However, both Meshed and Before still appear to be a valid options for reviewing material, this analysis only shows that Before is not significantly better than Meshed and vice versa. 

The Wilcoxon test is not without its weaknesses. Due to the large amount of ties within the data we are unaware of just how much it might have affected the results. 

The study did not make any observations with how the subjects felt about the method. One method might have caused more stress or uneasiness than the other. Further studies could potentialy look into which method provides the smoothest transition into a higher level classes.  

------------------

##### Source 

Friendly, M. and Franklin, P. (1980) Interactive presentation in multitrial free recall. Memory and Cognition 8 265–270 [Personal communication from M. Friendly].

##### References 

Fox, J. (2008) Applied Regression Analysis and Generalized Linear Models, Second Edition. Sage.

Fox, J. and Weisberg, S. (2011) An R Companion to Applied Regression, Second Edition, Sage.
